Metal Mixing and Ejection in Dwarf Galaxies is Dependent on Nucleosynthetic Source

### 2018-08-02 main.tex
Update on Overleaf.

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{species_bar.png}\\
\includegraphics[width=0.99\linewidth]{species_bar_ISM.png}
\caption{Top: The mass fraction of each metal species in the \textit{full simulation box} contained in the halo, gas in the galaxy disk, and stars. Bottom: The mass fraction of species within \textit{the disk} alone in each phase of the ISM (bottom). The leftmost bar in each plot shows the sum of all metals. \textbf{AE: Relabel y-axis, space in All Metals and C, larger figures}}
\label{fig:species_fractions}
\end{figure*}

@@

Unsurprisingly the diffuse HIM contains the most metal enriched gas, as it is comprised predominately of metal enriched SN ejecta. Clearly this leads to long power-law tails towards high metal fractions for each species, with a very narrow, poorly defined peak at low metal fractions. Generally, in colder gas the metal PDFs become less enriched with broader low metal fraction components and steeper, power-law tails. Although the extended power-law tail of the HIM leads to a large range of metal fractions, the HIM represents very little mass, and this tail represents recent, un-mixed enrichment. The comparatively narrow width in the log-normal regimes of CNM is perhaps surprising. Although this gas is at lower metal fraction than the WNM and WIM, it would appear that it is more well mixed. This runs counter to the idea that mixing times should be long in colder gas unless mixing first proceeds rapidly in the hot phases before mixing in with cold, disconnected structures across the galaxy.

@@

For O, the median mass fraction is ordered by phase temperature. The HIM is significantly more enriched than the cooler phases by anywhere from 0.1 dex to 4 dex, fluctuating by $\pm 1$ dex over the simulation time. The frequent large skew in the HIM (see second panel) and spread in the HIM (right two panels), coupled with its continual fluctuation, suggests that the HIM is not in equilibrium.
%This phase is poorly mixed because it is short lived, either cooling into different phases of the ISM or being ejected from the galaxy, and is continually enriched by new SN events.
Each cooler phase in O is progressively less enriched (lower median), with smaller skew and spread. The offset between phases and increasingly well mixed gas from hot to cold indicates that metal enrichment in the ISM of galaxies proceeds first through mixing on large scales in a hot phase, before progressively cooling through multiple phases until enriching star-forming gas. This 
%mm is the only explanation for [that's perhaps too strong a claim]
    can explain
how the cold gas can rapidly homogenize over the whole galaxy within $\sim$50 Myr, roughly when the cold phase exhibits a nearly constant spread (right panels). Individual enrichment events 
%mm will 
have significantly higher metallicities than the ambient ISM in any phase, and 
%mm will 
drive an increase in the difference between the mean and median of the PDF. These can be seen as the obvious spikes in the HIM and WIM. For O, the lack of these spikes in the two cold phases suggests again that enrichment does not occur directly in these phases, but proceeds more gradually through the warmer phases first.

@@

Although these trends are generally true for Ba, its evolution is much more complicated.
%In addition, the HIM is much less offset from the rest of the phases than is seen in O, with much less variability.
Unlike O, the spread and skew in Ba for the CNM, WNM, and WIM \textit{increase} in during the evolution, reaching differences in 90$^{\rm th}$ and 10$^{\rm th}$ percentiles of nearly 2 dex in the CNM and over 2 dex in the WNM and WIM. The HIM is seemingly unaffected by this trend, and simply fluctuates throughout the simulation. Given their lower energies, AGB winds more directly enrich the WNM and WIM, not the HIM as in SNe. The consequence of this is clear in Figure~\ref{fig:log-normal} by the wider PDFs and longer power-law tails in Ba in these phases. These tails represent the most recently enriched gas, 
%mm and 
        which
is clearly much more locally confined than O.

%{\bf (move to discussion?)}:
The positive skew in all of the PDFs presented here implies that most of the mass of the galaxy has a metal fraction below what one would normally adopt as the average metallicity (i.e.\ the ratio between the total mass of metals and the total mass). Assuming star-forming gas follows the same properties as the cold gas, this distinction is small ($\sim$ 0.2 dex), 
%mm but
   though 
significant, for elements released during SNe explosions, but can be very significant, up to $\sim$ 0.8 dex, for elements released in AGB winds. Chemical evolution models, especially one-zone models, follow the mean metal fraction, rather than the median. 
%mm These results indicate that these 
            Our results indicate that such
models are biased to overestimate gas and stellar elemental abundances.
. % However star formation may not be unbiased, and may preferentially come from gas sitting on the higher metal fraction tails.

To summarize, Figure~\ref{fig:phase-statistics} demonstrates: 1) there are qualitative differences in how SN injected elements (e.g. O, Mg) and AGB wind injected elements (e.g. N, Ba) are distributed through the ISM, with the latter having a broader range of variation and being less well mixed in all phases except the HIM; 2) hotter phases are more metal enriched, both because the cooler phases make up most of the initially unenriched mass of the ISM and because the hotter phases are more directly populated by recent enrichment events; 3) the cooler, denser phases, particularly for SN injected elements, are more well mixed than the hot phases of the ISM; 4) the PDFs of metal mass fraction are best fit by a log-normal + power law function; and therefore 5) the median metallicity available for star formation 
%mm rates 
lies below the mean galactic value. In the case of SN injected elements, enrichment proceeds quickly through the HIM over the entire galaxy. For AGB injected elements, enrichment proceeds through the WIM and WNM, leading to longer mixing timescales and larger metal fraction variations in the ISM.

\section{Discussion}
\label{sec:discussion}
We begin with a simple toy model that motivates the power-law tail at high metal fractions of the metal fraction PDFs and a subsequent turnover at low metal fractions in Section~\ref{sec:interpretation}. This work is placed in context with previous works focusing on metal mixing in the ISM in Section~\ref{sec:context}. We discuss the relevance of additional AGB yields not directly followed in this study in Section~\ref{sec:discussion:metal yields}. Finally, in Section~\ref{sec:stellar abundances} we discuss how these results relate to stellar abundances, make generalizations to more massive galaxies in Section~\ref{sec:massive galaxies}, and discuss possible impact of these results on chemical enrichment from more exotic nucleosynthetic sources in Section~\ref{sec:exotic enrichment}.

@@

In reality, however, $M_o$ is not an inexhaustible reservoir of primordial gas. Eventually the entire ambient medium will become enriched to some $Z > Z_o$, and $p(Z)$ will consist entirely of a truncated power-law. As enrichment proceeds, gas near the low metallicity truncation of $p(Z)$, which still comprises much of the mass of the system, enriches towards higher metallicities in the power-law tail. This will produce a turnover at the low $Z$ limit of $p(Z)$. The low-turnover limit is produced by diffusion from many different sources, and is thus multiplicative process, which, as we noted above, tends to produce log-normal distributions. 
%mm [additive vs multiplicative]
   %{\bf [According to Section~\ref{sec:log-normal}, MULTIPLICATIVE processes produce log-normals. Additive processes produce power-laws.  However, see Mouri (2013) for a possible counterexample...]}
The physical interpretation of these two components is that the power-law tail represents newly enriched and poorly mixed gas that is above the average gas metallicity and is undergoing dilution, while the log-normal component represents the ambient medium that lies below the average gas metallicity and is undergoing enrichment.

@@

\subsection{Impact on Stellar Abundance Patterns}
\label{sec:stellar abundances}
In order to better understand the physics driving stellar abundance patterns and distributions it is important to characterize the chemical evolution of star-forming gas in our simulations. This could be used to help disentangle sources of scatter in observed stellar abundance, including radial/azimuthal abundance gradients and stellar migration, redshift evolution, asymmetric accretion of pristine gas, and the intrinsic scatter in ISM abundances. If the 
%mm star-forming gas in our simulations can also be well-fit by a log-normal + power-law PDF
        metal mass fractions in star-forming gas in our simulations 
        can also be well fit by log-normal + power-law PDFs,
and if we can parameterize their evolution as functions of global galaxy properties, this 
%mm can 
    could
be used as a powerful 
%mm and analytic 
tool for modeling stellar abundance patterns in semi-analytic models. This would be a physically motivated way to account for both the intrinsic spread in stellar abundances due to inhomogeneities in the ISM and variations in mixing for different metal species. We reserve an analysis of these distributions in abundance space in both the gas and stars for future work. However, we briefly discuss the connection to star-forming gas and stellar enrichment below.

@@

\section{Conclusions}
\label{sec:conclusions}
%mm For the first time we 
We present a detailed analysis of galactic chemodynamics and metal mixing on an element-by-element basis in a low mass dwarf galaxy with hydrodynamics simulations that simultaneously capture multi-channel stellar feedback in detail with a multi-phase ISM. This high resolution simulation, coupled with our star-by-star modelling of stellar nucleosynthetic yields, has allowed us to analyze
%mm
      for the first time
how individual metal species couple to the ISM and galactic wind of this galaxy. We find that individual metal species do not share the same dynamical evolution, with differences directly related to nucleosynthetic origin (AGB winds, winds from massive stars, or core collapse SNe)
%mm , and Type Ia SNe).  [I'm inclined to leave them out as we really didn't model their influence]
This difference is most significant in our model between elements ejected predominately by AGB winds and those ejected predominantly by core-collapse SNe. In addition, we find 
%mm for the first time 
    the novel result
that the mass fraction PDFs of each metal in the ISM can be described using an analytic piecewise log-normal + power-law PDF. The properties of these PDFs vary with both ISM phase and metal species, again driven mostly by differences in enrichment sources.

@@

\item Hotter phases have metal fraction PDFs that are more enriched, with significant power-law tails as compared to cold phases, which have a more %mm dominant log-normal structure. 
   prominent log-normal component.
The lack of significant tails in the cold-phase PDFs indicates that metal mixing occurs rapidly in hotter phases before cooling into dense gas.

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{species_bar.png}\\
\includegraphics[width=0.99\linewidth]{species_bar_ISM.png}
\caption{Top: The mass fraction of each metal species in the \textit{full simulation box} contained in the halo, gas in the galaxy disk, and stars. Bottom: The mass fraction of species within \textit{the disk} alone in each phase of the ISM (bottom). The leftmost bar in each plot shows the sum of all metals.}
\label{fig:species_fractions}
\end{figure*}

@@

Unsurprisingly the diffuse HIM contains the most metal enriched gas, as it is comprised predominately of metal enriched SN ejecta. Clearly this leads to long power-law tails towards high metal fractions for each species, with a very narrow, poorly defined peak at low metal fractions. Generally, in colder gas the metal PDFs become less enriched with broader, low-metal-fraction components and steeper, power-law tails. Although the extended power-law tail of the HIM leads to a large range of metal fractions, the HIM represents very little mass, and this tail represents recent, un-mixed enrichment. The comparatively narrow width in the log-normal regimes of the CNM is perhaps surprising. Although this gas is at lower metal fraction than the WNM and WIM, it would appear that it is more well mixed. This runs counter to the idea that mixing times should be long in colder gas, unless (as we argue) mixing first proceeds rapidly in the hot phases before mixing in with cold, disconnected structures across the galaxy.

@@

For O, the median mass fraction is ordered by phase temperature. The HIM is significantly more enriched than the cooler phases by anywhere from 0.1 dex to 4 dex, fluctuating by $\pm 1$ dex over the simulation time. The frequent large skew in the HIM (see second panel) and spread in the HIM (right two panels), coupled with its continual fluctuation, suggests that the HIM is not in equilibrium.
%This phase is poorly mixed because it is short lived, either cooling into different phases of the ISM or being ejected from the galaxy, and is continually enriched by new SN events.
Each cooler phase in O is progressively less enriched (lower median), with smaller skew and spread. The offset between phases and increasingly well mixed gas from hot to cold indicates that metal enrichment in the ISM of galaxies proceeds first through mixing on large scales in a hot phase, before progressively cooling and/or mixing through multiple phases until enriching star-forming gas. This 
%mm is the only explanation for [that's perhaps too strong a claim]
    can explain
how the cold gas can rapidly homogenize over the whole galaxy within $\sim$50 Myr, roughly when the cold phase exhibits a nearly constant spread (right panels). Individual enrichment events 
%mm will 
have significantly higher metallicities than the ambient ISM in any phase, and 
%mm will 
drive an increase in the difference between the mean and median of the PDF. These can be seen as the obvious spikes in the HIM and WIM. For O, the lack of these spikes in the two cold phases suggests again that enrichment does not occur directly in these phases, but proceeds more gradually through the warmer phases first.

@@

Although these trends are generally true for Ba, its evolution is much more complicated.
%In addition, the HIM is much less offset from the rest of the phases than is seen in O, with much less variability.
Unlike O, the spread and skew in Ba for the CNM, WNM, and WIM \textit{increase} during the evolution, reaching differences in 90$^{\rm th}$ and 10$^{\rm th}$ percentiles of nearly 2 dex in the CNM and over 2 dex in the WNM and WIM. The HIM is seemingly unaffected by this trend, and simply fluctuates throughout the simulation. Given their lower energies, AGB winds more directly enrich the WNM and WIM, not the HIM as in SNe. The consequence of this is clear in Figure~\ref{fig:log-normal} by the wider PDFs and longer power-law tails in Ba in these phases. These tails represent the most recently enriched gas, 
%mm and 
        which
is clearly much more locally confined than O.

%{\bf (move to discussion?)}:
The positive skew in all of the PDFs presented here implies that most of the mass of the galaxy has a metal fraction below what one would normally adopt as the average metallicity (i.e.\ the ratio between the total mass of metals and the total mass). Assuming star-forming gas follows the same properties as the cold gas, this distinction is small ($\sim$ 0.2 dex), 
%mm but
   though 
significant, for elements released during SNe explosions, but can be very significant, up to $\sim$ 0.8 dex, for elements released in AGB winds. Chemical evolution models, especially one-zone models, follow the mean metal fraction, rather than the median. 
%mm These results indicate that these 
            Our results indicate that such
models are biased to overestimate gas and stellar elemental abundances.
% However star formation may not be unbiased, and may preferentially come from gas sitting on the higher metal fraction tails.

To summarize, Figure~\ref{fig:phase-statistics} demonstrates: 1) there are qualitative differences in how SN-injected elements (e.g. O, Mg) and AGB-wind-injected elements (e.g. N, Ba) are distributed through the ISM, with the latter having a broader range of variation and being less well mixed in all phases except the HIM; 2) hotter phases are more metal enriched, both because the cooler phases make up most of the initially unenriched mass of the ISM and because the hotter phases are more directly populated by recent enrichment events; 3) the cooler, denser phases, particularly for SN injected elements, are more well mixed than the hot phases of the ISM; 4) the PDFs of metal mass fraction are best fit by a log-normal + power law function; and therefore 5) the median metallicity available for star formation 
%mm rates 
lies below the mean galactic value. In the case of SN injected elements, enrichment proceeds quickly through the HIM over the entire galaxy. For AGB injected elements, enrichment proceeds through the WIM and WNM, leading to longer mixing timescales and larger metal fraction variations in the ISM.

\section{Discussion}
\label{sec:discussion}
We begin with a simple toy model that motivates the power-law tail at high metal fractions of the metal fraction PDFs and a subsequent turnover at low metal fractions in Section~\ref{sec:interpretation}. This work is placed in context with previous papers focusing on metal mixing in the ISM in Section~\ref{sec:context}. We discuss the relevance of additional AGB yields not directly followed in this study in Section~\ref{sec:discussion:metal yields}. Finally, in Section~\ref{sec:stellar abundances} we discuss how these results relate to stellar abundances, make generalizations to more massive galaxies in Section~\ref{sec:massive galaxies}, and discuss possible impacts of these results on chemical enrichment from more exotic nucleosynthetic sources in Section~\ref{sec:exotic enrichment}.

@@

In reality, however, $M_o$ is not an inexhaustible reservoir of primordial gas. Eventually the entire ambient medium will become enriched to some $Z > Z_o$, and $p(Z)$ will consist entirely of a truncated power-law. As enrichment proceeds, gas near the low metallicity truncation of $p(Z)$, which still comprises much of the mass of the system, enriches towards higher metallicities in the power-law tail. This will produce a turnover at the low $Z$ limit of $p(Z)$. The low-turnover limit is produced by diffusion from many different sources, and is thus a multiplicative process, which, as we noted above, tends to produce log-normal distributions. 
%mm [additive vs multiplicative]
   %{\bf [According to Section~\ref{sec:log-normal}, MULTIPLICATIVE processes produce log-normals. Additive processes produce power-laws.  However, see Mouri (2013) for a possible counterexample...]}
The physical interpretation of these two components is that the power-law tail represents newly enriched and poorly mixed gas that is above the average gas metallicity and is undergoing dilution, while the log-normal component represents the ambient medium that lies below the average gas metallicity and is undergoing enrichment.

@@

\subsection{Impact on Stellar Abundance Patterns}
\label{sec:stellar abundances}
In order to better understand the physics driving stellar abundance patterns and distributions it is important to characterize the chemical evolution of star-forming gas in our simulations. This could be used to help disentangle sources of scatter in observed stellar abundances, including radial/azimuthal abundance gradients and stellar migration, redshift evolution, asymmetric accretion of pristine gas, and the intrinsic scatter in ISM abundances. If the 
%mm star-forming gas in our simulations can also be well-fit by a log-normal + power-law PDF
        metal mass fractions in star-forming gas in our simulations 
        can also be well fit by log-normal + power-law PDFs,
and if we can parameterize their evolution as functions of global galaxy properties, this 
%mm can 
    could
be used as a powerful 
%mm and analytic 
tool for modeling stellar abundance patterns in semi-analytic models. This would be a physically motivated way to account for both the intrinsic spread in stellar abundances due to inhomogeneities in the ISM and variations in mixing for different metal species. We reserve an analysis of these distributions in abundance space in both the gas and stars for future work. However, we briefly discuss the connection to star-forming gas and stellar enrichment below.

@@

\section{Conclusions}
\label{sec:conclusions}
%mm For the first time we 
We present a detailed analysis of galactic chemodynamics and metal mixing on an element-by-element basis in a low mass dwarf galaxy with hydrodynamics simulations that simultaneously capture multi-channel stellar feedback in detail with a multi-phase ISM. This high resolution simulation, coupled with our star-by-star modelling of stellar nucleosynthetic yields, has allowed us to analyze
%mm
      for the first time
how individual metal species couple to the ISM and galactic wind of this galaxy.

We find that individual metal species do not share the same dynamical evolution, with differences directly related to nucleosynthetic origin (AGB winds, winds from massive stars, or core collapse SNe)
%mm , and Type Ia SNe).  [I'm inclined to leave them out as we really didn't model their influence]
This difference is most significant, in our model, between elements ejected predominately by AGB winds and those ejected predominantly by core-collapse SNe. In addition, we find 
%mm for the first time 
    the novel result
that the mass fraction PDFs of each metal in the ISM can be described using an analytic piecewise log-normal + power-law PDF. The properties of these PDFs vary with both ISM phase and metal species, again driven mostly by differences in enrichment sources.

@@

\item Hotter phases have metal fraction PDFs that are more enriched, with significant power-law tails as compared to cold phases, which have a more %mm dominant log-normal structure. 
   prominent log-normal component.
The lack of significant tails in the cold-phase PDFs indicates that metal mixing occurs rapidly in hotter phases before cooling and/or mixing into denser gas.
